---
title: "Data Wrangling Exercise 2"
subject: "Dealing wtih Missing Values"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1: Port of embarkation
The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.)

```{r cars}
library(readxl)
library(xlsx)
library(dplyr)
library(tidyr)
lapply("titanic3.xls", function(f) {
  df = read.xlsx(f, sheetName = "titanic3")
  write.csv(df, gsub(".xls","_original.csv", f), row.names=FALSE)
})
titanic <- read.csv("titanic3_original.csv", header = TRUE, sep = ",")

dsc <- tbl_df(titanic)
```

## Part 1b

Port of embarkation variable: find the missing values and replace them with S.

```{r pressure, echo=FALSE}

is.data.frame(dsc)
## we want if there are any values that appear in the data to be na 
any(is.na(dsc$embarked))
## asking if any of variables contain nas if so how much
sapply(dsc, function(x) sum(is.na(x)))

dsc$embarked[is.na(dsc$embarked)] <- "S"
## to check if there is any missing values
head(dsc$embarked)
```

## Part 2: Replace missing age values with the mean age

```{r}
dsc$age[is.na(dsc$age)] <- mean(dsc$age,na.rm=TRUE)
## double check if there are any missing. Since it ran succesfully, then it should come as FAlSE once you run the code down below.
any(is.na(dsc$age))
```

## Why would pick mean over the other solutions?

Think about other ways you could have populated the missing values in the age column. Why would you pick any of those over the mean (or not)?

Answer:# We could also have grouped the data by gender and calculated mean and median if there was a correlation
 However, doing this through the code altered_data  %>% group_by(sex) %>% summarise(mean(age)) shows that there is a small difference, but not significant. Short of getting more data on missing values (is there a correlation between missing age and other factors?), this is a good approximation
 
#3: Lifeboat
You’re interested in looking at the distribution of passengers in different lifeboats, but as we know, many passengers did not make it to a boat :-( 
This means that there are a lot of missing values in the boat column. Fill these empty slots with a dummy value e.g. the string 'None' or 'NA'
 
```{r}
class(dsc$boat)
## since it is  factor we change to character
dsc$boat <- as.character(dsc$boat)
dsc$boat[is.na(dsc$boat)] <- "NA"


```
## Cabin Variable
 the cabin number is missing might be a useful indicator of survival, thus create a new column has_cabin_number which has 1 if there is a cabin number and 0 other wise.
 
```{r}
df <- dsc %>% mutate(has_cabin_number = ifelse(cabin != "",1,0))
```

There is a clustering of cabins alphetically- hence it is because a roll call was done alphabetically for the first lifeboats, increasing the odds of surviving for people with lucky names
```{r}
titanic.final <- write.csv(df, file="titanic.clean.csv")
```


