---
title: "Data Wrangling Exercise 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data and Clean up brand names

0: Load the data in RStudio
Save the data set as a CSV file called refine_original.csv and load it in RStudio into a data frame.
1: Clean up brand names
Clean up the 'company' column, so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).


```{r cars}
require(dplyr)
require(tidyr)
require(xlsx)
require(dummies)
refine <- read.csv("refine_original.csv")

refine_c <- tbl_df(refine)

```

## Including Plots

1: Clean up the brand names
Clean up the 'company' column, so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

```{r pressure, echo=FALSE}
refine_c<- refine_c %>%
  arrange(refine_c)  %>%
  mutate_each(funs(tolower), matches("company"))
```

Changing typos found on the data manually

```{r}
refine_c$company <- sub(".*ps","philips",refine_c$company, ignore.case = TRUE)
refine_c$company<- sub(".*ak.*","akzo",refine_c$company, ignore.case = TRUE)
refine_c$companyy <- sub(".*ten","van houten",refine_c$company, ignore.case = TRUE)
refine_c$company <- sub(".*ver","unilever",refine_c$company, ignore.case = TRUE)
```

Seperate product and code number using tidyr 
```{r}
refine_c <- separate(refine_c,"Product.code...number", c("product_code", "product_number"), sep = "-")
```

```{r}
product <- c("Smartphone", "TV", "Laptop", "Tablet")
product_code <- c("p", "v", "x", "q")
product_df <- data.frame(product, product_code)
refine_c <- left_join(refine_c, product_df, by="product_code")
```


4: Add full address for geocoding
You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.


```{r}
refine_c <- refine_c %>% unite(full_address, address, city, country, sep=",")
```

5: Create dummy variables for company and product category
Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.
Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever
Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet
```{r}
refine_c$company_philips <- as.numeric(refine_c$company %in% "philips")
refine_c$company_akzo <- as.numeric(refine_c$company %in% "akzo")
refine_c$company_vanhouten <- as.numeric(refine_c$company %in% "van houten")
refine_c$company_unilever <- as.numeric(refine_c$company %in% "unilever")
```



```{r}
refine_dummy <- refine_c %>%
  mutate(company_akzo = ifelse(company == "akzo", 1, 0), company_philips = ifelse(company == "philips", 1,0), company_unilever = ifelse(company == "unilever",1,0), company_van_houten = ifelse(company == "van houten", 1,0)) %>% 
  mutate(product_smartphone = ifelse(product == "Smartphone",1,0), product_tablet = ifelse(product == "Tablet",1,0), product_tv = ifelse(product == "TV",1,0), product_laptop = ifelse(product == "Laptop",1,0))
```

Write the data in cleaned up csv file.

```{r}
refine_final <- write.csv(refine_dummy, "refine_clean.csv", row.names=FALSE)
```

